---
title: "Generic Data Generation"
author: "Gigi"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: true
    df_paged: paged
---
# Public liability

The aim of this document is to help make data generation as easy as possible. I will be looking at public liability insurance only in this document.

By using a set of generic functions hopefully it can be relatively easy to build up a complete data set from scratch.

# Setup

```{r setup-hidden, include=FALSE}
rm( list = ls() )

knitr::opts_chunk$set(
    echo = TRUE,
    fig.height = 8,
    fig.width = 11,
    cache = FALSE
)

set.seed( 107 )
```

We use the packages:

* `tidyverse` is used to manipulate and visualise data.
* `lubridate` is used to convert and manipulate dates.

```{r setup-library}
library( tidyverse )
library( lubridate )
```

There are also functions specific for the data generation, they are defined in an external file. 

```{r setup-source}
source( 'lib.R' )
# source( 'utility.R' )
```

## Setting initial parameters

```{r set-general-parameters}
policy_count <- 1000
incident_base_rate <- 0.5
```

## Reading in tables

Set any required variables for reading in the tables.

```{r set-tbl-options}
# Define basic column data types
col_types = cols(
    out_value = col_character(),
    odds      = col_double()
)
```

### Stand alone tables

Read in the locations table.

```{r read-in-locations}
postcode_tbl <- read_csv(
    'data_out/locations_tbl.csv',
    progress = FALSE,
    col_types = cols(
        postcode  = col_character(),
        latitude  = col_double(),
        longitude = col_double()
    )
)
```

Read in the table stipulating the maximum number of days in each month.

```{r read-in-max-days}
max_days_tbl <- read_csv(
    'data_in/max_days_per_month.csv',
    progress = FALSE,
    col_types = col_types
) %>% 
    calcCumulative
```

### Policies tables

```{r read-in-tbls-policies}
policy_years_odds_tbl <- read_csv(
    'data_in/policy_years.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_months_odds_tbl <- read_csv(
    'data_in/policy_months.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_length_odds_tbl <- read_csv(
    'data_in/policy_length.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_size_rating_odds_tbl <- read_csv(
    'data_in/policy_size_rating.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_claim_limit_odds_tbl <- read_csv(
    'data_in/claim_limit.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_no_claims_odds_tbl <- read_csv(
    'data_in/no_claims_no.csv',
    col_types = col_types
) %>% 
    calcCumulative

size_rating_numeric_tbl <- read_csv(
    'data_in/size_rating_numeric.csv',
    col_types = cols(
        filter_1 = col_character(),
        min = col_double(),
        max = col_double(),
        type = col_character(),
        mean = col_double(),
        sd = col_double(),
        df = col_integer()
    )
)

claim_limit_tbl <- read_csv(
    'data_in/claim_limit_extra.csv',
    col_types = cols(
        filter_1 = col_integer(),
        premium_rate = col_double()
    )
)

no_claims_tbl <- read_csv(
    'data_in/no_claims_no_extra.csv',
    col_types = cols(
        filter_1      = col_integer(),
        discount_rate = col_double()
    )
)
```

### Incidents tables

```{r read-in-tbls-incidents}
# date related tables
years_odds_tbl <- read_csv( 
    'data_in/years.csv', 
    col_types = col_types 
) %>% 
    calcCumulative

months_odds_tbl <- read_csv( 
    'data_in/months.csv',
    col_types = col_types
) %>% 
    calcCumulative
```

# Building the tables

## Policies table

There are many features required for the policies table. 

* ID.
* Various rating factors such as the number of employees, turnover, number of years no claims.
* Some charged premium.
 
```{r generate-policies-tbl}  
policies_tbl <- policy_count %>% 
    # extract a postcode for each policy
    sample_n( tbl = postcode_tbl, ., replace = TRUE ) %>% 
    select( postcode ) %>% 
    # add in the start year and month
    mutate( 
        start_year  = level1Values( policy_years_odds_tbl, n() ),
        start_month = level1Values( policy_months_odds_tbl, n() )
    ) %>% 
    # add in the start_day, this relies on the start_month
    group_by( start_month ) %>% 
    mutate(
        start_day = sample_n( 
            tibble(
                val = seq( 
                    1,  
                    filterTbl( max_days_tbl, start_month[1] ) %>% 
                        as.numeric
                )
            ),
            n(),
            replace = TRUE
        ) %>% 
            .[[1]]
    ) %>% 
    ungroup %>% 
    # create a single start date, then calc the end date based 
    # on the policy length
    mutate(
        start_date    = paste( start_day, start_month, start_year, sep = '-' ) %>% 
            dmy,
        policy_length = level1Values( policy_length_odds_tbl, n() ) %>% 
            as.numeric,
        end_date      = start_date %m+% 
            months( policy_length )
    ) %>% 
    # assign a size rating, I want these distributed roughly the same over
    # each year
    group_by( start_year ) %>% 
    mutate(
        size_rating = level1Values( policy_size_rating_odds_tbl, n() )
    ) %>% 
    # estimate an 'employee count' rating factor by the size rating
    group_by( size_rating ) %>% 
    mutate(
        employee_count = numericValues(
            n    = n(),
            type = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                filter_1_val = size_rating[1],
                select_col   = 'type'
            ),
            mean = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                filter_1_val = size_rating[1],
                select_col   = 'mean'
            ) %>% 
                as.numeric,
            minimum = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                filter_1_val = size_rating[1],
                select_col   = 'min'
            ) %>% 
                as.numeric,
            maximum = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                filter_1_val = size_rating[1],
                select_col   = 'max'
            ) %>% 
                as.numeric,
            sd = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                filter_1_val = size_rating[1],
                select_col   = 'sd'
            ) %>% 
                as.numeric,
            df = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                filter_1_val = size_rating[1],
                select_col   = 'df'
            ) %>% 
                as.numeric
        )
    ) %>% 
    ungroup %>% 
    mutate(
        claim_limit = level1Values( policy_claim_limit_odds_tbl, n() ) %>% 
            as.numeric,
        no_claims_years = level1Values( policy_no_claims_odds_tbl, n() ) %>% 
            as.numeric
    ) %>% 
    group_by( no_claims_years ) %>% 
    mutate(
        no_claims_discount = filterTbl( 
            no_claims_tbl, 
            no_claims_years[1],
            'discount_rate'
        ) %>% 
            as.numeric
    ) %>% 
    group_by( claim_limit ) %>% 
    mutate(
        premium_rate = filterTbl( 
            tbl          = claim_limit_tbl, 
            filter_1_val = claim_limit[1],
            select_col   = 'premium_rate'
        ) %>% 
            as.numeric
    ) %>% 
    ungroup %>% 
    mutate(
        premium   = (
            premium_rate * claim_limit * 
            log( employee_count ) * 
            ( 1 + no_claims_discount )
        ) %>% 
            round( 2 ),
        incidents = incident_base_rate * employee_count * 
            numericValues( 
                n(), 
                'chisq', 
                df = 0.1, 
                minimum = 0.5,
                maximum = 3.5
            )
    )
```

## Incidents table

```{r generate-incidents-tbl}
incidents_tbl <- policies_tbl[ 
    rep( rownames( policies_tbl ) %>% as.integer, policies_tbl$incidents ), 
] %>% 
    # Remove intermediate columns
    select( 
        postcode, start_date, end_date, claim_limit
    ) %>% 
    mutate(
        policy_days = start_date %--%
            end_date %>% 
            as.duration,
        policy_days = policy_days / ddays( 1 ),
        incident_date = ( start_date + ddays( runif( n(), 0, policy_days ) ) ) %>% 
            date
    )
```