---
title: "Pulic Liability Data Generation"
author: "Gigi"
date: "`r Sys.Date()`"
output: 
    html_document: 
        toc: yes
        toc_float: true
        df_paged: paged
---
# Public liability

The aim of this document is to help make data generation as easy as possible. I will be looking at public liability insurance only in this document.

By using a set of generic functions hopefully it can be relatively easy to build up a complete data set from scratch.

# Setup

```{r setup-hidden, include=FALSE}
rm( list = ls() )

knitr::opts_chunk$set(
    echo       = TRUE,
    fig.height = 8,
    fig.width  = 11,
    cache      = FALSE
)

set.seed( 185 )
```

We use the packages:

* `tidyverse` is used to manipulate and visualise data.
* `lubridate` is used to convert and manipulate dates.
* `stringr` is used to manipulate strings, it is part of the tidyverse.

```{r setup-library}
library( tidyverse )
library( lubridate )
library( stringr )
```

There are also functions specific for the data generation, they are defined in an external file. 

```{r setup-source}
source( 'lib.R' )
```

```{r setup-to-remove, include=FALSE}
function_list_to_remove <- c(
    "calcCumulative", "filterTbl", "level1AdjValues", "levelValues", 
    "numericValues"
)
```

## Setting initial parameters

```{r set-general-parameters}
policy_count <- 5000
incident_base_rate <- 0.5
date_recieved <- dmy( '01-06-2016' )
```

```{r set-file-out-names}
policies_out_name  <- 'data_out/policies.csv'
incidents_out_name <- 'data_out/incidents.csv'
claims_out_name    <- 'data_out/claims.csv'
```

```{r set-general-hidden, include=FALSE}
general_list_to_remove <- c(
    "policy_count", "incident_base_rate", "col_types", 
    "policies_out_name", "incidents_out_name", "claims_out_name",
    "date_recieved"
)
```

## Reading in tables

Set any required variables for reading in the tables.

```{r set-tbl-options}
# Define basic column data types
col_types = cols(
    out_value = col_character(),
    odds      = col_double()
)
```

### Stand alone tables

Read in the locations table.

```{r read-in-locations}
postcode_tbl <- read_csv(
    'data_out/locations_tbl.csv',
    progress = FALSE,
    col_types = cols(
        postcode  = col_character(),
        latitude  = col_double(),
        longitude = col_double()
    )
)
```

Read in the table stipulating the maximum number of days in each month.

```{r read-in-max-days}
max_days_tbl <- read_csv(
    'data_in/public_liability/max_days_per_month.csv',
    progress = FALSE,
    col_types = col_types
) %>% 
    calcCumulative
```

### Policies tables

```{r read-in-tbls-policies}
policy_years_odds_tbl <- read_csv(
    'data_in/public_liability/policy_years.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_months_odds_tbl <- read_csv(
    'data_in/public_liability/policy_months.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_length_odds_tbl <- read_csv(
    'data_in/public_liability/policy_length.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_size_rating_odds_tbl <- read_csv(
    'data_in/public_liability/policy_size_rating.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_claim_limit_odds_tbl <- read_csv(
    'data_in/public_liability/claim_limit.csv',
    col_types = col_types
) %>% 
    calcCumulative

policy_no_claims_odds_tbl <- read_csv(
    'data_in/public_liability/no_claims_no.csv',
    col_types = col_types
) %>% 
    calcCumulative

size_rating_numeric_tbl <- read_csv(
    'data_in/public_liability/size_rating_numeric.csv',
    col_types = cols(
        filter_1 = col_character(),
        min = col_double(),
        max = col_double(),
        type = col_character(),
        mean = col_double(),
        sd = col_double(),
        df = col_integer()
    )
)

claim_limit_tbl <- read_csv(
    'data_in/public_liability/claim_limit_extra.csv',
    col_types = cols(
        filter_1 = col_integer(),
        premium_rate = col_double()
    )
)

no_claims_tbl <- read_csv(
    'data_in/public_liability/no_claims_no_extra.csv',
    col_types = cols(
        filter_1      = col_integer(),
        discount_rate = col_double()
    )
)
```

### Incidents tables

```{r read-in-tbls-incidents}
years_odds_tbl <- read_csv( 
    'data_in/public_liability/years.csv', 
    col_types = col_types 
) %>% 
    calcCumulative

months_odds_tbl <- read_csv( 
    'data_in/public_liability/months.csv',
    col_types = col_types
) %>% 
    calcCumulative

incident_cat_1 <- read_csv(
    'data_in/public_liability/incident_category_1.csv',
    col_types = col_types
) %>% 
    calcCumulative

incident_cat_1_adj <- read_csv(
    'data_in/public_liability/incident_category_1_adj.csv',
    col_types = col_types
) %>% 
    calcCumulative

incident_cat_2 <- read_csv(
    'data_in/public_liability/incident_category_2.csv',
    col_types = col_types
) %>% 
    calcCumulative

claim_flag <- read_csv(
    'data_in/public_liability/claim_flag.csv',
    col_types = col_types
) %>% 
    calcCumulative
```

### Claims tables

```{r read-in-tbls-claims}
claim_reported_days <- read_csv(
    'data_in/public_liability/claim_reported_day_adj.csv',
    col_types = col_types
)

claim_closed_days <- read_csv(
    'data_in/public_liability/claim_closed_days.csv',
    col_types = col_types
)

claim_status_tbl <- read_csv(
    'data_in/public_liability/claim_status.csv',
    col_types = col_types
) %>% 
    calcCumulative

claim_incurred <- read_csv(
    'data_in/public_liability/claim_incurred.csv',
    col_types = col_types
) %>% 
    calcCumulative

claim_paid <- read_csv(
    'data_in/public_liability/claim_paid.csv',
    col_types = col_types
)
```

```{r read-in-hidden, include=FALSE}
# These are the odds tables that won't be required at the end
table_list_to_remove <- c(
    "postcode_tbl", "max_days_tbl", "policy_years_odds_tbl", 
    "policy_months_odds_tbl", "policy_length_odds_tbl", 
    "policy_size_rating_odds_tbl", "policy_claim_limit_odds_tbl",
    "policy_no_claims_odds_tbl", "size_rating_numeric_tbl",
    "claim_limit_tbl", "no_claims_tbl", "years_odds_tbl",
    "months_odds_tbl", "incident_cat_1", "incident_cat_2",
    "incident_cat_1_adj", "claim_flag", "claim_reported_days",
    "claim_closed_days", "claim_status_tbl", "claim_incurred",
    "claim_paid"
)
```

# Building the tables

## Policies table

There are many features required for the policies table. 

* ID.
* Various rating factors such as the number of employees, turnover, number of years no claims.
* Some charged premium.
 
```{r generate-policies-tbl}  
policies_tbl <- policy_count %>% 
    # extract a postcode for each policy
    sample_n( tbl = postcode_tbl, ., replace = TRUE ) %>% 
    # add in the start year and month
    mutate( 
        start_year  = levelValues( policy_years_odds_tbl, n() ),
        start_month = levelValues( policy_months_odds_tbl, n() )
    ) %>% 
    # add in the start_day, this relies on the start_month
    group_by( start_month ) %>% 
    mutate(
        start_day = sample_n( 
            tibble(
                val = seq( 
                    1,  
                    filterTbl( max_days_tbl, start_month[1] ) %>% 
                        as.numeric
                )
            ),
            n(),
            replace = TRUE
        ) %>% 
            .[[1]]
    ) %>% 
    ungroup %>% 
    # create a single start date, then calc the end date based 
    # on the policy length
    mutate(
        start_date    = paste( start_day, start_month, start_year, sep = '-' ) %>% 
            dmy,
        policy_length = levelValues( policy_length_odds_tbl, n() ) %>% 
            as.numeric,
        end_date      = start_date %m+% 
            months( policy_length )
    ) %>% 
    # assign a size rating, I want these distributed roughly the same over
    # each year
    group_by( start_year ) %>% 
    mutate(
        size_rating = levelValues( policy_size_rating_odds_tbl, n() )
    ) %>% 
    # estimate an 'employee count' rating factor by the size rating
    group_by( size_rating ) %>% 
    mutate(
        employee_count = numericValues(
            n    = n(),
            type = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                size_rating[1],
                select_col   = 'type'
            ),
            mean = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                size_rating[1],
                select_col   = 'mean'
            ) %>% 
                as.numeric,
            minimum = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                size_rating[1],
                select_col   = 'min'
            ) %>% 
                as.numeric,
            maximum = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                size_rating[1],
                select_col   = 'max'
            ) %>% 
                as.numeric,
            sd = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                size_rating[1],
                select_col   = 'sd'
            ) %>% 
                as.numeric,
            df = filterTbl( 
                tbl          = size_rating_numeric_tbl,
                size_rating[1],
                select_col   = 'df'
            ) %>% 
                as.numeric
        )
    ) %>% 
    ungroup %>% 
    mutate(
        claim_limit     = levelValues( policy_claim_limit_odds_tbl, n() ) %>% 
            as.numeric,
        no_claims_years = levelValues( policy_no_claims_odds_tbl, n() ) %>% 
            as.numeric
    ) %>% 
    group_by( no_claims_years ) %>% 
    mutate(
        no_claims_discount = filterTbl( 
            no_claims_tbl, 
            no_claims_years[1],
            select_col = 'discount_rate'
        ) %>% 
            as.numeric
    ) %>% 
    group_by( claim_limit ) %>% 
    mutate(
        premium_rate = filterTbl( 
            claim_limit_tbl, 
            claim_limit[1],
            select_col   = 'premium_rate'
        ) %>% 
            as.numeric
    ) %>% 
    ungroup %>% 
    mutate(
        premium   = (
            premium_rate * claim_limit * 
            log( employee_count ) * 
            ( 1 - no_claims_discount )
        ) %>% 
            round( 2 ),
        incidents = incident_base_rate * employee_count * 
            numericValues( 
                n(), 
                'chisq', 
                df = 0.1, 
                minimum = 0.5,
                maximum = 3.5
            ) *
            round( policy_length / 12, 0 ),
        policy_id = paste0( 
            "PL", start_year, size_rating, 
            str_pad( string = row_number(), width = 7, pad = "0" )
        )
    )
```

## Incidents table

```{r generate-incidents-tbl}
incidents_tbl <- policies_tbl[ 
    rep( rownames( policies_tbl ) %>% as.integer, policies_tbl$incidents ), 
] %>% 
    # Remove intermediate columns
    select( 
        policy_id, postcode, start_date, end_date, start_year, claim_limit,
        latitude, longitude
    ) %>% 
    mutate(
        policy_days = start_date %--%
            end_date %>% 
            as.duration,
        policy_days = policy_days / ddays( 1 ),
        incident_date = ( start_date + ddays( runif( n(), 0, policy_days ) ) ) %>% 
            date
    ) %>% 
    group_by( start_year ) %>% 
    mutate(
        description_level_1 = levelValues( tbl = incident_cat_1, n = n() ),
        description_level_1 = level1AdjValues(
            org_values       = description_level_1,
            filter_value     = start_year,
            modification_tbl = incident_cat_1_adj,
            n                = n()
        )
    ) %>% 
    group_by( description_level_1 ) %>% 
    mutate(
        description_level_2 = levelValues(
            tbl     = incident_cat_2,
            description_level_1[1],
            n            = n()
        )
    ) %>% 
    ungroup %>% 
    mutate(
        incident_year = year( incident_date )
    ) %>% 
    group_by( incident_year ) %>% 
    mutate(
        claim_flag = levelValues(
            tbl     = claim_flag,
            incident_year,
            n            = n()
        ) %>% 
            as.logical
    ) %>% 
    ungroup
```

## Claims table

```{r generate-claim-tbl}
claims_tbl <- incidents_tbl %>% 
    filter( claim_flag ) %>% 
    select( -claim_flag ) %>% 
    group_by( description_level_1, description_level_2 ) %>% 
    mutate(
        claim_reported_days = numericValues( 
            n = n(),
            type = filterTbl(
                claim_reported_days,
                description_level_2[1], description_level_1[1],
                select_col = "type", default = "norm"
            ),
            df = filterTbl(
                claim_reported_days,
                description_level_2[1], description_level_1[1],
                select_col = "df", default = 4, convert = "num"
            ),
            multiplier = filterTbl(
                claim_reported_days,
                description_level_2[1], description_level_1[1],
                select_col = "multiplier", default = 1, convert = "num"
            ),
            addition = filterTbl(
                claim_reported_days,
                description_level_2[1], description_level_1[1],
                select_col = "addition", default = 0, convert = "num"
            ),
            min = filterTbl(
                claim_reported_days,
                description_level_2[1], description_level_1[1],
                select_col = "min", default = -Inf, convert = "num"
            ),
            max = filterTbl(
                claim_reported_days,
                description_level_2[1], description_level_1[1],
                select_col = "max", default = Inf, convert = "num"
            )
        ),
        claim_reported_date = incident_date + ddays( claim_reported_days )
    ) %>% 
    group_by( description_level_2 ) %>% 
    mutate(
        claim_closed_days = numericValues(
            n = n(),
            type = filterTbl(
                claim_closed_days,
                description_level_2[1],
                select_col = 'type', default = 'norm'
            ),
            df = filterTbl(
                claim_closed_days,
                description_level_2[1],
                select_col = 'df', default = 4, convert = 'num'
            ),
            addition = filterTbl(
                claim_closed_days,
                description_level_2[1],
                select_col = 'addition', default = 0, convert = 'num'
            ),
            shape = filterTbl(
                claim_closed_days,
                description_level_2[1],
                select_col = 'shape', default = 0, convert = 'num'
            ),
            rate = filterTbl(
                claim_closed_days,
                description_level_2[1],
                select_col = 'rate', default = 0, convert = 'num'
            )
        ),
        claim_closed_date = claim_reported_date + ddays( claim_closed_days )
    ) %>% 
    ungroup %>% 
    mutate(
        claim_reported_date = if_else(
            claim_reported_date %--% date_recieved <= 0,
            dmy( NA ),
            claim_reported_date
        ),
        claim_closed_date = if_else(
            claim_closed_date %--% date_recieved <= 0,
            dmy( NA ),
            claim_closed_date
        )
    ) %>% 
    group_by( description_level_2 ) %>% 
    mutate(
        claim_status = ifelse(
            is.na( claim_closed_date %--% date_recieved ),
            levelValues(
                claim_status_tbl,
                n(),
                description_level_2,
                default = "Open"
            ),
            "Closed"
        )
    ) %>% 
    group_by( incident_year ) %>% 
    mutate(
        incurred_amt_tmp = levelValues(
            claim_incurred,
            n(),
            incident_year,
            default = "1"
        )
    ) %>% 
    group_by( incident_year, incurred_amt_tmp ) %>% 
    mutate(
        incurred_amt = numericValues(
            n(),
            type = filterTbl(
                claim_incurred %>% rename( filter_2 = out_value ),
                incident_year, incurred_amt_tmp,
                select_col = "type"
            ),
            df = filterTbl(
                claim_incurred %>% rename( filter_2 = out_value ),
                incident_year, incurred_amt_tmp,
                select_col = "df", convert = "num"
            ),
            value = filterTbl(
                claim_incurred %>% rename( filter_2 = out_value ),
                incident_year, incurred_amt_tmp,
                select_col = "value", convert = "num"
            ),
            multiplier = filterTbl(
                claim_incurred %>% rename( filter_2 = out_value ),
                incident_year, incurred_amt_tmp,
                select_col = "multiplier", convert = "num"
            ),
            round = 2
        )
    ) %>% 
    group_by( incident_year ) %>% 
    mutate(
        claim_paid_tmp = numericValues(
            n(),
            type = filterTbl(
                claim_paid,
                incident_year,
                select_col = "type"
            ),
            value = filterTbl(
                claim_paid,
                incident_year,
                select_col = "value",
                convert = "num"
            ),
            mean = filterTbl(
                claim_paid,
                incident_year,
                select_col = "mean",
                convert = "num"
            ),
            sd = filterTbl(
                claim_paid,
                incident_year,
                select_col = "sd",
                convert = "num"
            ),
            minimum = filterTbl(
                claim_paid,
                incident_year,
                select_col = "minimum",
                convert = "num"
            ),
            maximum = filterTbl(
                claim_paid,
                incident_year,
                select_col = "maximum",
                convert = "num"
            ),
            round = 4
        )
    ) %>% 
    ungroup %>% 
    mutate(
        claim_paid = round( incurred_amt * claim_paid_tmp, 2 ),
        claim_reserved = incurred_amt - claim_paid
    ) %>%
    select( -incurred_amt_tmp, -claim_paid_tmp )
```

# Write files out

```{r write-out}
write_csv( policies_tbl, policies_out_name )
write_csv( incidents_tbl, incidents_out_name )
write_csv( claims_tbl, claims_out_name )
```

# Clean up

After the generation remove all the tables, functions, and values that were used for the generation. Leaving only the generated tables as outputs.

```{r clean-up}
rm( list = table_list_to_remove )
rm( list = function_list_to_remove )
rm( list = general_list_to_remove )
rm( list = c(
    "table_list_to_remove",
    "function_list_to_remove",
    "general_list_to_remove"
    )
)
```